---
title: "Who will vote for Trump & Biden?"
author: "RuiLiang&rl2886"
date: "9/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
library(haven)
library(tidyverse)
library(naniar)
library(labelled)
library(ggplot2)
```

## Load data

```{r echo=FALSE}
pilot_data <- read_sav("/Users/raymondliang/Documents/GitHub/Fall2020-Project1-rl2886-ray/data/anes_pilot_2020ets_sav/anes_pilot_2020ets_sav.SAV")
```

## Missing data checking
Column vidx has two missing data.Vidx it's the question about 'dose R know State Voter ID law?',therefore to be fair, I would eliminate this two rows of data.
```{r echo=FALSE}
any(is.na(pilot_data))
pilot_data %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
pilot_data_1<-pilot_data[-which(is.na(pilot_data$vidx)),]
```

## Population of the data 
The survey sample is consistent with the race demographic data from US census bureau, which mean our inference from data will be more precise.

```{r echo=FALSE}
asian<-sum(pilot_data$race1_asian)
black<-sum(pilot_data$race1_black)
hpi<-sum(pilot_data$race1_hpi)
namer<-sum(pilot_data$race1_namer)
white<-sum(pilot_data$race1_white)

barplot(c(asian,black,hpi,namer,white),names=c('asian','black','hpi','namer','white'))
```
party ID
```{r}

ggplot(as.data.frame(pilot_data$pid1r), aes(x=as.factor(pilot_data$pid1r) )) +
  geom_bar( )


```

religious
```{r}

ggplot(as.data.frame(pilot_data$relig1), aes(x=as.factor(pilot_data$relig1) )) +
  geom_bar( )


```

income
```{r}

ggplot(as.data.frame(pilot_data$inc_anes), aes(x=as.factor(pilot_data$inc_anes) )) +
  geom_bar( )

```

exp
```{r}
# travel exp
ggplot(as.data.frame(pilot_data$exptravel_ever), aes(x=as.factor(pilot_data$exptravel_ever) )) +
  geom_bar( )

```


demographics
```{r}
# birth
ggplot(as.data.frame(pilot_data$birthyr), aes(x=as.factor(pilot_data$birthyr) )) +
  geom_bar( )
mean(pilot_data$birthyr)
#sex
ggplot(as.data.frame(pilot_data$sex), aes(x=as.factor(pilot_data$sex) )) +
  geom_bar( )
#educ
ggplot(as.data.frame(pilot_data$educ), aes(x=as.factor(pilot_data$educ) )) +
  geom_bar( )
#children
ggplot(as.data.frame(pilot_data$children), aes(x=as.factor(pilot_data$children) )) +
  geom_bar( )
```

## Feeling towards President
Under our data,the feeling rate towards president Trump  is  44.59 which means he is not favorable by our sample.In comparison, last president Obama is 59.25 and 2020 president candidate Joe Biden is 51.67.It seems like although the feelings toward Joe Biden is not a great favorable feeling, his rate still trump President trump.

```{r echo=FALSE}
mean(pilot_data$fttrump1)
mean(pilot_data$ftobama1)
mean(pilot_data$ftbiden1)
```

##How does each races feeling towards these three person?

```{r echo=FALSE}
mean(pilot_data$fttrump1[pilot_data$race1_black==1])
mean(pilot_data$fttrump1[pilot_data$race1_asian==1])
mean(pilot_data$fttrump1[pilot_data$race1_white==1])
mean(pilot_data$fttrump1[pilot_data$race1_hpi==1])
mean(pilot_data$fttrump1[pilot_data$race1_namer==1])

mean(pilot_data$ftobama1[pilot_data$race1_black==1])
mean(pilot_data$ftobama1[pilot_data$race1_asian==1])
mean(pilot_data$ftobama1[pilot_data$race1_white==1])
mean(pilot_data$ftobama1[pilot_data$race1_hpi==1])
mean(pilot_data$ftobama1[pilot_data$race1_namer==1])

mean(pilot_data$ftbiden1[pilot_data$race1_black==1])
mean(pilot_data$ftbiden1[pilot_data$race1_asian==1])
mean(pilot_data$ftbiden1[pilot_data$race1_white==1])
mean(pilot_data$ftbiden1[pilot_data$race1_hpi==1])
mean(pilot_data$ftbiden1[pilot_data$race1_namer==1])

```

## How does people think about President Trump's job performance?
People are diverse towards the performance of Trump's.The number of people who approve Trump is 1474 which is larger than 1404, the number of people who disapprove Trump.

```{r echo=FALSE}
ggplot(as.data.frame(pilot_data$apppres7), aes(x=as.factor(pilot_data$apppres7) )) +
  geom_bar( )
length(pilot_data$apppres7[pilot_data$apppres7<4])
length(pilot_data$apppres7[pilot_data$apppres7>4])
```

Eco performance
Overall, people think he did a good job in Eco

```{r echo=FALSE}
ggplot(as.data.frame(pilot_data$econpres7), aes(x=as.factor(pilot_data$econpres7) )) +
  geom_bar( )
```

Covid performance
There are a lot of NA data, which doesnt make sense.
```{r echo=FALSE}
ggplot(as.data.frame(pilot_data$covidpres7), aes(x=as.factor(pilot_data$covidpres7) )) +
  geom_bar( )
pilot_data$mis_covid1
ggplot(as.data.frame(pilot_data$mis_covid19), aes(x=as.factor(pilot_data$mis_covid19) )) +
  geom_bar( )
```

## Voting
The number of people who would like to vote Joe biden is lightly over Donald Trump.What kind of people would JB or Trump?
```{r echo=FALSE}
d1<-pilot_data$vote20jb

ggplot(as.data.frame(d1), aes(x=as.factor(d1) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )
```

## Trait
Traits describe Donald Trump
```{r echo=FALSE}
d2<-pilot_data$dtleader1
ggplot(as.data.frame(d2), aes(x=as.factor(d2) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

d3<-pilot_data$dtleader2
ggplot(as.data.frame(d3), aes(x=as.factor(d3) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )
     
d4<-pilot_data$dtcares
ggplot(as.data.frame(d4), aes(x=as.factor(d4) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

d5<-pilot_data$dthonest
ggplot(as.data.frame(d5), aes(x=as.factor(d5) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

d6<-pilot_data$dtdiv
ggplot(as.data.frame(d6), aes(x=as.factor(d6) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

d7<-pilot_data$dtknow
ggplot(as.data.frame(d7), aes(x=as.factor(d7) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )
```
Traits describe Joe Biden
```{r echo=FALSE}
j2<-pilot_data$jbleader1
ggplot(as.data.frame(j2), aes(x=as.factor(j2) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

j3<-pilot_data$jbleader2
ggplot(as.data.frame(j3), aes(x=as.factor(j3) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )
     
j4<-pilot_data$jbcares
ggplot(as.data.frame(j4), aes(x=as.factor(j4) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

j5<-pilot_data$jbhonest
ggplot(as.data.frame(j5), aes(x=as.factor(j5) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

j6<-pilot_data$jbdiv
ggplot(as.data.frame(j6), aes(x=as.factor(j6) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )

j7<-pilot_data$jbknow
ggplot(as.data.frame(j7), aes(x=as.factor(j7) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7,0.9) )
```
Faith in experts
```{r}

ggplot(as.data.frame(pilot_data$experts), aes(x=as.factor(pilot_data$experts) )) +
  geom_bar( )

ggplot(as.data.frame(pilot_data$science), aes(x=as.factor(pilot_data$science) )) +
  geom_bar( )

ggplot(as.data.frame(pilot_data$exphelp), aes(x=as.factor(pilot_data$exphelp) )) +
  geom_bar( )
```
##Other reference:
https://www.pewresearch.org/politics/2020/08/13/perceptions-of-trump-and-biden/
